---
title: "The Clock Catalog"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup, include = FALSE}
library(methylCIPHERv2)
```

```{r, echo = FALSE}
catalog <- list_clocks(all_columns = TRUE)

shown <- c(
  "clock_id", "group_id", "n_cpgs", "covariates", "normalize",
  "tissues_derived", "array_type_trained", "training_algorithm",
  "cohort_trained", "n_samples_trained", "health_status_trained",
  "age_min_trained", "age_max_trained", "age_unit_trained",
  "sex_distribution_trained", "ancestry_trained"
)

reactable::reactable(
  catalog[shown],
  searchable = TRUE,
  filterable = TRUE,
  striped = TRUE,
  highlight = TRUE,
  compact = TRUE,
  wrap = FALSE,
  resizable = TRUE,
  defaultPageSize = 25,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(25, 50, 100),
  defaultSorted = "group_id",
  defaultColDef = reactable::colDef(minWidth = 140),
  # the description is a paragraph, so it expands under the row instead of
  # taking a column of its own
  details = function(index) {
    text <- catalog[["description"]][[index]]
    if (is.na(text)) {
      text <- "No description on record for this clock."
    }
    htmltools::div(style = "padding: 1rem; max-width: 60rem;", text)
  },
  columns = list(
    clock_id = reactable::colDef(name = "Clock", minWidth = 200),
    group_id = reactable::colDef(name = "Group", minWidth = 160),
    n_cpgs = reactable::colDef(name = "CpGs", minWidth = 90),
    covariates = reactable::colDef(name = "Covariates", minWidth = 110),
    normalize = reactable::colDef(name = "Normalize", minWidth = 110),
    tissues_derived = reactable::colDef(name = "Tissue", minWidth = 150),
    array_type_trained = reactable::colDef(name = "Array", minWidth = 130),
    training_algorithm = reactable::colDef(name = "Algorithm", minWidth = 170),
    cohort_trained = reactable::colDef(name = "Cohort", minWidth = 260),
    n_samples_trained = reactable::colDef(name = "Samples", minWidth = 100),
    health_status_trained = reactable::colDef(name = "Health", minWidth = 150),
    age_min_trained = reactable::colDef(name = "Age min", minWidth = 100),
    age_max_trained = reactable::colDef(name = "Age max", minWidth = 100),
    age_unit_trained = reactable::colDef(name = "Age unit", minWidth = 120),
    sex_distribution_trained = reactable::colDef(name = "Sex", minWidth = 130),
    ancestry_trained = reactable::colDef(name = "Ancestry", minWidth = 200)
  )
)
```

## What the columns mean

The table scrolls sideways to reach the columns that do not fit. Select the arrow at the start of a row to read a longer description of that clock.

**Clock** is the identifier for a single scored clock. **Group** is the family it belongs to. **CpGs** counts the CpGs that the clock scores.

**Covariates** lists the columns that `pheno` must carry for that clock. An empty cell means the clock reads methylation values alone.

**Normalize** names the background normalization scheme a clock declares, and an empty cell means it declares none. A `quantile` cell is on by default and a `bmiq` cell is off. The `normalize` argument of `calc_clocks()` turns either one on or off. Normalization also needs a background panel that is much larger than the scoring panel, so `calc_clocks()` warns and declines a scheme whose background is too thin to use.

The remaining columns describe the clock as it was published. **Tissue**, **Array**, **Cohort**, **Samples**, **Health**, **Age min**, **Age max**, **Age unit**, **Sex** and **Ancestry** report the data that the clock was trained on. **Algorithm** names the method that fit it. An empty cell means the value is not on record.

`codebook()` returns the same descriptions as a data frame, for a set of clocks or for the clocks that a run scored.

```{r}
codebook(c("Horvath1", "Hannum"))[["column"]]
```

## Scoring from the catalog

Any value in the **Clock** column works as the `clocks` argument.

```{r, eval = FALSE}
res <- calc_clocks(DNAm, clocks = "Horvath1", pheno = pheno)
```

A group name scores every member of that family.

```{r, eval = FALSE}
res <- calc_clocks(DNAm, clocks = "GrimAge", pheno = pheno)
```

A tag scores every clock that carries it. `list_clock_tags()` returns the tags with their members.

```{r}
names(list_clock_tags())
```

## Filtering the catalog in code

`list_clocks()` returns this catalog as a data frame, so it is available in a script as well as on this page. It takes a tag directly.

```{r}
mitotic <- list_clocks(tag = "mitotic")
mitotic[["clock_id"]]
```

The result is an ordinary data frame, so you can filter the other columns with base R.

```{r}
all_clocks <- list_clocks()
bundled <- all_clocks[!all_clocks[["external"]], ]
nrow(bundled)
```
